---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OBIShmpr

<!-- badges: start -->
<!-- badges: end -->

The goal of OBIShmpr is to link and query OBIS occurence data with a variety of marine habitat data.

## Installation

<!-- You can install the released version of OBIShmpr from [CRAN](https://CRAN.R-project.org) with: 

``` r
install.packages("OBIShmpr")
```
-->
You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("annakrystalli/OBIShmpr")
```
## Example

### Get habitat data for a species

```{r example}
library(OBIShmpr)
## basic example code
```


## An example with one species
This is an example of how to run the above code for a single species - we use _Scytothamnus fasciculatus_, Aphia ID 325567, chosen as it has just 6 OBIS records so should run reasonably quickly.

```{r, message = FALSE, warning  = FALSE, eval = FALSE}
get_temp_summ_by_sp(sp_id = 325567)
```




### Get species profile for habitat
