---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OBIShmpr

<!-- badges: start -->
<!-- badges: end -->

The goal of OBIShmpr is to link and query OBIS occurence data with a variety of marine habitat data.

### Data Sources

- #### [**Bio-ORACLE**](https://www.bio-oracle.org/) & [**MARSPEC**](http://www.marspec.org/) data accesed through  `sdmpredictors`:
    a set of GIS rasters providing geophysical, biotic and environmental data for surface and benthic marine realms. Accessed via [`sdmpredictors`](https://github.com/lifewatch/sdmpredictors)
    
- #### `sedmaps` [Synthetic map of the NW European
Shelf sedimentary environment](https://www.earth-syst-sci-data.net/10/109/2018/essd-10-109-2018.pdf)
    **data product** [10.15129/1e27b806-1eae-494d-83b5-a5f4792c46fc](https://pureportal.strath.ac.uk/en/datasets/data-for-a-synthetic-map-of-the-northwest-european-shelf-sediment)
    0.125◦ by 0.125◦ resolution synthetic maps of continuous properties of the north-west European sedimentary environment. The maps are a blend of gridded survey
data, statistically modelled values based on distributions of bed shear stress due to tidal currents and waves and bathymetric properties. 
    - percentage compositions of mud, sand and gravel;
    - porosity and permeability; 
    - median grain size of the whole sediment and of the sand and the gravel fractions;
    - carbon and nitrogen content of sediments; 
    - percentage of seabed area covered by rock; 
    - mean and maximum depth-averaged tidal velocity and wave orbital velocity at the seabed; and mean monthly natural disturbance rates. 
    
- #### [EMODnet Seabed Habitat maps]()
  

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("annakrystalli/OBIShmpr")
```
## Example

### Get habitat data for a species

```{r example}
library(OBIShmpr)
```


## An example with one species
This is an example of how to run the above code for a single species - we use _Scytothamnus fasciculatus_, Aphia ID 325567, chosen as it has just 6 OBIS records so should run reasonably quickly.

Choose some layers to extract data from

```{r}
sp_id <- 325567
layers <- c("BO2_nitratemax_bdmin", "BO_parmax", "BO2_tempmax_bdmax", "permeability", 
"tidal_vel_max", "april")
```

```{r, message = FALSE, warning  = FALSE, eval = TRUE}
obis_match_habitat(sp_id, layers = layers)
```



```{r}
layer_codes <- c("BO_ph", "BO_phosphate", 
                 "BO2_phosphatemean_bdmax", 
                 "BO_nitrate", "BO2_nitratemean_bdmax", 
                 "surface_nitrogen", "tn", "surface_carbon",
                 "BO_sstmean", "BO_sstmin", "BO_sstmax", 
                 "MS_biogeo13_sst_mean_5m", "MS_biogeo14_sst_min_5m",
                 "MS_biogeo15_sst_max_5m")
```



### _Palinurus elephas_

Next I use the AphiaID for _Palinurus elephas_.


```{r}
species_id <- worrms::wm_name2id("Palinurus elephas")

data <- OBIShmpr::obis_match_habitat(
    sp_id = species_id, 
    layers = layer_codes,
    geometry = bbox_nwes)

data
```

```{r}
library(ggplot2)
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")


ggplot() +
    geom_sf(data = world) +
    geom_sf(data = data, aes(color = BO_ph) ) +
    coord_sf(crs = 3035,
             xlim = c(2426378.0132, 7093974.6215),
             ylim = c(1308101.2618, 5446513.5222)) +
    theme_bw() +
    ggtitle("Palinurus elephas",
            subtitle = paste0('AphiaID = ', species_id))

```

