---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OBIShmpr

<!-- badges: start -->
<!-- badges: end -->

The goal of OBIShmpr is to link and query OBIS occurence data with a variety of marine habitat data.

### Data Sources

- #### [**Bio-ORACLE**](https://www.bio-oracle.org/) & [**MARSPEC**](http://www.marspec.org/) data accesed through  `sdmpredictors`:
    a set of GIS rasters providing geophysical, biotic and environmental data for surface and benthic marine realms. Accessed via [`sdmpredictors`](https://github.com/lifewatch/sdmpredictors)
    
- #### `sedmaps` [Synthetic map of the NW European
Shelf sedimentary environment](https://www.earth-syst-sci-data.net/10/109/2018/essd-10-109-2018.pdf)
    **data product** [10.15129/1e27b806-1eae-494d-83b5-a5f4792c46fc](https://pureportal.strath.ac.uk/en/datasets/data-for-a-synthetic-map-of-the-northwest-european-shelf-sediment)
    0.125◦ by 0.125◦ resolution synthetic maps of continuous properties of the north-west European sedimentary environment. The maps are a blend of gridded survey
data, statistically modelled values based on distributions of bed shear stress due to tidal currents and waves and bathymetric properties. 
    - percentage compositions of mud, sand and gravel;
    - porosity and permeability; 
    - median grain size of the whole sediment and of the sand and the gravel fractions;
    - carbon and nitrogen content of sediments; 
    - percentage of seabed area covered by rock; 
    - mean and maximum depth-averaged tidal velocity and wave orbital velocity at the seabed; and mean monthly natural disturbance rates. 
    
- #### [EMODnet Seabed Habitat maps]()
  

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("annakrystalli/OBIShmpr")
```
## Example

### Get habitat data for a species

```{r example}
library(OBIShmpr)
```


## An example with one species
This is an example of how to run the above code for a single species - we use _Scytothamnus fasciculatus_, Aphia ID 325567, chosen as it has just 6 OBIS records so should run reasonably quickly.

Choose some layers to extract data from

```{r}
sp_id <- 325567
layers <- c("BO2_nitratemax_bdmin", "BO_parmax", "BO2_tempmax_bdmax", "permeability", 
"tidal_vel_max", "april")
```

```{r, message = FALSE, warning  = FALSE, eval = TRUE}
obis_match_habitat(sp_id, layers = layers)
```




### Get species profile for habitat

## EMODnet Seabed Habitat Data

```{r}
get_emodnet_wfs_layers(layers = c("be000142", "be000144"))
```

